### 2020-11-17 index.Rmd
Hello world

---
title: "Everest Bio"
description: |
  Projet de recherche CREIDD et ERPI
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages

```