### 2020-11-17 slides/index.Rmd
adding ppt

---
title: "Everest Bio"
subtitle: |
  **ÉV**aluation du r**É**seau de **RE**cyclage pour **S**outenir la **T**ransition vers une **BIO**économie <br/>
  Projet de recherche CREIDD et ERPI
author: "Fabio Xie"
institute: "CREIDD, ERPI"
date: "2020/11/18 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

## The problem: Plastic waste

.pull-left[ 
- US recycled less than 9% of its plastic waste.
- EU, recycled about 30%. However, this number is what is collected, not from what is in circulation*. 
- China’s ban on plastic trash
- Plastic trash could top 13 billion tons by 2050
- Economics of plastic recycling historically have been challenging
- Little financial incentive to encourage consumers to recycle their own plastic waste

## We barely recycle plastic!!.. 
]

.pull-right[ 
![Plastic waste](https://images.unsplash.com/photo-1537084642907-629340c7e59c?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1567&q=80)
]

.footnote[
*Kranzinger, L., Pomberger, R., Schwabl, D., Flachberger, H., Bauer, M., Lehner, M., Hofer, W., 2018. Output-oriented analysis of the wet mechanical processing of polyolefin-rich waste for feedstock recycling. Waste Manag. Res. 36, 445–453
]

---

# Challengue

The development of sustainable industrial supply chains

.pull-left[
.small[

- Les seuls critères technico-économiques ne permettent pas (plus) d’avoir une prise de décision éclairée

- Besoin d’évaluer les filières de manière différente, notamment les impacts réels sur les écosystèmes

- Une vision plus holistique est nécessaire pour évaluer la faisabilité de ces nouvelles filières
]
 ]

---