### 2020-11-17 docs/slides/index.Rmd
version complete

---
title: "Everest Bio"
subtitle: |
  **ÉV**aluation du r**É**seau de **RE**cyclage pour **S**outenir la **T**ransition vers une **BIO**économie <br/>
  Projet de recherche CREIDD et ERPI
author: "Fabio Cruz / Mauricio Camargo"
institute: "CREIDD, ERPI"
date: "2020/11/18 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "assets/Fabio.css", "assets/Fabio-fonts.css"]
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "assets/macros.js"
---

```{r share-again, echo=FALSE}
xaringanExtra::use_share_again()
```

## The problem: Plastic waste

.pull-left[ 
- US recycled less than 9% of its plastic waste.
- EU, recycled about 30%. However, this number is what is collected, not from what is in circulation*. 
- China’s ban on plastic trash
- Plastic trash could top 13 billion tons by 2050
- Economics of plastic recycling historically have been challenging
- Little financial incentive to encourage consumers to recycle their own plastic waste

## We barely recycle plastic!!.. 
]

.pull-right[ 
![Plastic waste](https://images.unsplash.com/photo-1537084642907-629340c7e59c?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1567&q=80)
]

.footnote[
*Kranzinger, L., Pomberger, R., Schwabl, D., Flachberger, H., Bauer, M., Lehner, M., Hofer, W., 2018. Output-oriented analysis of the wet mechanical processing of polyolefin-rich waste for feedstock recycling. Waste Manag. Res. 36, 445–453
]

---

# Challengue

The development of sustainable industrial supply chains

.left-column[
.small[

- Only technico-economic criteria alone do not (no longer) enables a informed decision making

- Need to assess value chains in a different way, including actual impacts on ecosystems

- A more holistic vision is needed to assess the feasibility of these new sectors.
]
 ]

.right-column[ 
.polaroid[ ![](figures/Recycling.png) ]
]

---

# Towards a local distributed recycling approach

.pull-left[
- Local closed loop
- AM adds value
- Small quantities 
- Reduction of landfilling
- Minimize the use of virgin materials

![](Figures/Green-fablab-1.jpg)

]

.pull-right[

![](http://lf2l.fr/assets/images/3-projects/gf/Logo-GF.png)
http://lf2l.fr/projects/green-fablab

.center[![:scale 80%](Figures/Green-fablab-circuit.png)]

]

---

## Distributed recycling via additive manufacturing (DRAM)
Technical system

.center[![:scale 75%](Figures/Green-fablab-0.png)]

.footnote[
Cruz Sanchez, F.A., Boudaoud, H., Camargo, M., Pearce, J.M., 2020. Plastic recycling in additive manufacturing: A systematic literature review and opportunities for the circular economy. J. Clean. Prod. 264, 121602. https://doi.org/10.1016/j.jclepro.2020.121602 
]

---

## Distributed recycling via additive manufacturing (DRAM)
Technical system --> Global system

.center[![:scale 75%](Figures/DRAM-05.png)]

---

## Distributed recycling via additive manufacturing (DRAM)

Taking into account the environment and their ecosystem services (ES).

.pull-left[
![:scale 90%](Figures/Eco-Services-00.jpg)
]

--

.pull-right[
![](Figures/Eco-Services-01.png)

- How the future recycling chain could be including the environment?
]

.footnote[
Bruel, A., Kronenberg, J., Troussier, N., & Guillaume, B. (2019). Linking industrial ecology and ecological economics: A theoretical and empirical foundation for the circular economy. Journal of Industrial Ecology, 23(1), 12-21.
]

---

# Main Goals

- To develop a generic methodology to assess the ecosystem services rendered by industrial activities operating in short circuit in order to improve decision-making by industrial and public actors.

- Validation on the case of a short-circuit plastic recycling chain.

---

# Methods

- In line with the thesis work previously carried out by the two teams, development of a coupled Systems Dynamics and LCA approach.

- Application on a case study, the GreenFablab, by evaluating its environmental impact when setting up a short plastic recycling circuit.

.center[ ![:scale 80%](Figures/Goals.png)]

---

# Distributed recycling via additive manufacturing (DRAM)

Global system

.pull-left[
![](Figures/DRAM-07.png)
]

.pull-right[
- Establish and evaluate scenarios in a systemic and dynamic way

- Evaluate the influences of the main operating variables on the system

- Taking into account the priorities of the multiple stakeholders

- Facilitate your decision making public-private actors
]
.footnote[
Santander, P., Cruz Sanchez, F.A., Boudaoud, H., Camargo, M., 2020. Closed loop supply chain network for local and distributed plastic recycling for 3D printing: a MILP-based optimization approach. Resour. Conserv. Recycl. 154, 104531. https://doi.org/10.1016/j.resconrec.2019.104531 
]

---

# Delivrables

1. An article in an indexed international journal
2. Participation in an international conference 
3. An operational model with a model of a software tool for assessing local impacts
4. Final report on the evaluation of Green Fablab and possible generalisations for short industrial circuits.

---

# Valorisations

- Method and prototype of a tool for the evaluation of local impacts of circular production systems

---
class: middle, center

Thanks !

![:scale 70%](Figures/DRAM-bd.png)