### 2020-11-20 00-References.Rmd
adding refs

---
title: "References"
description: |
  Some additional details about the website
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
repository_url: https://github.com/LF2L/everest-bio
#bibliography: Biblio.bib
#nocite: '@*'
---

# Circular Economy

## On the definition

- Calisto Friant, M., Vermeulen, W.J.V., Salomone, R., 2020. A typology of circular economy discourses: Navigating the diverse visions of a contested paradigm. Resour. Conserv. Recycl. 161, 104917. https://doi.org/10.1016/j.resconrec.2020.104917

- Morseletto, P., 2020. Targets for a circular economy. Resour. Conserv. Recycl. 153, 104553. https://doi.org/10.1016/j.resconrec.2019.104553

- Schöggl, J.-P., Stumpf, L., Baumgartner, R.J., 2020. The narrative of sustainability and circular economy - A longitudinal review of two decades of research. Resour. Conserv. Recycl. 163, 105073. https://doi.org/10.1016/j.resconrec.2020.105073

# Service Ecosystémiques

## 2016

- Clement Morlat. Modélisation dynamique des systèmes de coûts pour une gestion durable des territoires. Economies et finances. Université Paris-Saclay, 2016. Français. ffNNT : 2016SACLV115ff.
fftel-01617502f. https://tel.archives-ouvertes.fr/tel-01617502/document

# Developpement durable

##

---

Coming soon more ! `r emo::ji("smile")`