### 2020-11-27 03-comportements.Rmd
Rationale for article 3

---
title: "Sustainable behaviour"
description: |
  some
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
repository_url: https://github.com/LF2L/everest-bio
bibliography: 03-comportement.bib
nocite: '@*'

author:
  - first_name: " Master 3"
    last_name: ""
    url: 
    affiliation: CREIDD?, ERPI?
    affiliation_url:

  - first_name: "Fabio"
    last_name: "Cruz"
    url: https://erpi.univ-lorraine.fr/people/Fabio-Cruz/
    affiliation: ERPI
    affiliation_url: https://erpi.univ-lorraine.fr/
    orcid_id: 0000-0001-8529-5327

  - first_name: "Nadège"
    last_name: "Troussier"
    url: https://recherche.utt.fr/icd-directory/nadege-troussier
    affiliation: CREIDD
    affiliation_url: https://recherche.utt.fr/icd-directory/nadege-troussier
    orcid_id: 0000-0001-6707-0300

  - first_name: "Hakim"
    last_name: "Boudaoud"
    url: https://erpi.univ-lorraine.fr/people/Hakim-Boudaoud/
    affiliation: ERPI
    affiliation_url: https://erpi.univ-lorraine.fr/
    orcid_id: 0000-0003-4452-5952

  - first_name: "Mauricio"
    last_name: "Camargo"
    url: https://erpi.univ-lorraine.fr/people/Mauricio-Camargo/
    affiliation: ERPI
    affiliation_url: https://erpi.univ-lorraine.fr/
    orcid_id: 0000-0003-3867-2438

---

# Abstract

Master 3: Mise en place de la collecte des bouchons : pourrait être analyse sur le comportement des acteurs => quel changement dans ces profils de gens ? => autre master (identifier ce que ça change chez les gens qui participent ou pas) avec Tatiana et Laure !!! doctorant avec master côté Nancy (avec Hakim et Mauricio)

# Rationale

Some ideas:

- Circular economy decisions are very complex but often the perceptions, attitudes, and personal values of decision makers directly influence organizations’ strategic actions towards a circular economy. Both active participation of organizations and the role of decision makers are pivotal to implement a circular economy for plastics [@Khan2020].

- Identifying and assessing specific barriers associated with the workplace context could be crucial for achieving better employee engagement and a lower ecological footprint for organizations [@Yuriev2020]