### 2021-10-25 slides/partials/EU-projects.Rmd

---
class: middle, left
exclude: true

background-image: url(`r paste0(URL, 'Climatelabs/Climatelabs.png')`)
background-position: center
background-size: contain

---
class: middle
exclude: true

# EU projects

.pull-left[
```{r , include=TRUE, echo=FALSE, fig.cap=" ", out.width = '60%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Smagrinet/Smagrinet-00.jpg'))
```
https://www.smagrinet.eu/

```{r , include=TRUE, echo=FALSE, fig.cap=" ", out.width = '60%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Digiterri/Digiterri'))
```

https://digiterri.eu/

] 
.pull-right[

```{r , include=TRUE, echo=FALSE, fig.cap=" ", out.width = '100%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Inedit/Inedit-00.png'))
```

https://www.inedit-project.eu/
]