### 2021-11-09 xar-EB-general/index.Rmd

---
title: "Everest Bio"
subtitle: |
  **ÉV**aluation du r**É**seau de **RE**cyclage pour **S**outenir la **T**ransition vers une **BIO**économie <br/>
  Projet de recherche CREIDD et ERPI
author: "Fabio Cruz / Mauricio Camargo / Nadège Troussier"
institute: "CREIDD, ERPI"
date: '2020/11/18 (updated: `r Sys.Date()`)'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default"]
    seal: false 
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "assets/macros.js"
      navigation:
        scroll: false # disable slide transitions by scrolling
---

```{r setup, include = FALSE, echo=FALSE, message=FALSE}
options(htmltools.dir.version = FALSE)

knitr::opts_chunk$set(include= TRUE, echo=FALSE, message=FALSE, fig.path="figures/")

library(readxl)
library(here)
library(flextable)
library(tidyverse)

# Link for the Figures
URL = c('https://raw.githubusercontent.com/fabbiocrux/Figures/main/')

# CICES File
#file_cices <- list.files("../data/CICES")
CICES <- read_excel(here("../data/CICES/Finalised-V5.1_18032018.xlsx"), sheet = "CICES V5.1",  skip = 4 )
```

```{r, include=TRUE, echo=FALSE}
# CSS for the CSS
htmltools::includeCSS("https://raw.githubusercontent.com/fabbiocrux/presentaciones/gh-pages/Assets-global/css/Fabio-fonts.css")
htmltools::includeCSS("https://raw.githubusercontent.com/fabbiocrux/presentaciones/gh-pages/Assets-global/css/Fabio.css")
```

```{r share-again, echo=FALSE}
xaringanExtra::use_share_again()
```

---

class: title-slide, left, middle
background-image: url(`r paste0(URL, 'Everest-bio/logos/Logos.png')`)
background-position: 0px 0px
background-size: contain

# `r rmarkdown::metadata$title`

## `r rmarkdown::metadata$subtitle`

### `r rmarkdown::metadata$author`

### `r rmarkdown::metadata$date`

---

```{r, include=FALSE}
Fabio <- TRUE
```

```{r, child= if (!Fabio) 'partials/Fabio-presentation.Rmd'}

```

---

class: middle

# Agenda

.pull-left[
1. .content-box-yellow[ERPI Laboratory]
1. Green Fablab
1. Everest Bio
1. Proposition and Research axis
]

.pull-right[

.brand-blue[Equipe de Recherche sur le Processus Innovatifs - ERPI]

```{r, out.width = '80%', fig.align='center'}
knitr::include_graphics('https://erpi.univ-lorraine.fr/assets/images/logo-ERPI.svg')
```
]

---

## .small[Equipe de Recherche sur le Processus Innovatifs - ERPI]

A .red[Multidisciplinary] team in **Industrial Engineering** specialized in the study of Innovation processes

.pull-left[

```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics('https://erpi.univ-lorraine.fr/assets/images/logo-ERPI.svg')
```

45 Members average

24 permanent researchers <br>
(10 full Pr., 13 associated Pr, 1 Research engineer) <br>
1 Studies engineer <br>
5 Postdoc  <br>
3 Temporary lecturers  <br>
2 Admin. <br>

]

.pull-right[

```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'LF2L/Methodology-complet-00.jpg'))
```

Research platform: **LF2L** - http://lf2l.fr
]

---

# Research domains at ERPI:

.pull-left[
.pull-left[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/01-Innovation-metrology.jpg'))
```

.orange[
**Innovation capability**

Product <br>
Process  <br>
Business Model <br>

]
]
.pull-right[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/02-Prospective.jpg'))
```
.blue[
**Technological  Emergence/maturity**

Product-Process <br>
Industrial supply chain <br>
]]]

.pull-right[
.pull-left[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/03-Design.jpg'))
```
.green[
**Materialized**

Creative Exploration <br>
Feasibility Product-Process <br>
Physic & digital (virtual) Prototyping <br>
]
]
.pull-right[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/04-Acceptability.jpg'))
```
.purple[
**Front-end validation**
Multi-actors Acceptability <br>
User eXperience & test by use <br>
Living Lab <br>
]]]

.center.Large[Creator of **methods & tools** supporting innovation process]

---

## Scope: The Fuzzy Front End of innovation

.pull-left[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/Scope-00.jpg'))
```
]

.pull-right[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/Purpose-00.jpg'))
```
.footnote[
Martinez-Hernandez, E., 2017. Trends in sustainable process design—from molecular to global scales. Curr. Opin. Chem. Eng. 17, 35–41. <br/> https://doi.org/10.1016/j.coche.2017.05.005 
]
]

Think earlier in the sustainable design and their impact

---

## Innovation: a response to current social challenges

```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/Fields-00.jpg'))
```

.tiny[Tran Thi Hoang, G., Dupont, L., & Camargo, M. (2019). Application of Decision-Making Methods in Smart City Projects: A Systematic Literature Review. Smart Cities, 2(3), 433-452.]

---

class: middle

# Agenda

.pull-left[
1. Who am I?
2. ERPI Laboratory
3. .content-box-yellow[Green Fablab]
4. Everest Bio
5. Proposition and Research axis
]

.pull-right[

```{r, out.width = '100%', fig.align='center'}
knitr::include_graphics('http://lf2l.fr/assets/images/3-projects/gf/Logo-GF.png')
```
*Circular economy and 3D printing: Towards a distributed recycling paradigm*
]

---

class: top, left
background-image: url(`r paste0(URL, 'Plastics/Plastic-problem-03.svg')`)
background-position: right
background-size: 70%

## Plastics waste since  olden times

.left-column[
.tiny[
- US recycled less than 9% of its plastic waste.
- EU, recycled about 30%. However, this number is what is collected, not from what is in circulation*. 
- China’s ban on plastic trash
- Plastic trash could top 13 billion tons by 2050
- Economics of plastic recycling historically have been challenging
- Little financial incentive to encourage consumers to recycle their own plastic waste

```{r, out.width = '60%', fig.align='center'}
knitr::include_graphics('https://images.unsplash.com/photo-1537084642907-629340c7e59c?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1567&q=80')
```

]]

.footnote[
Geyer, R., Jambeck, J.R., Law, K.L., 2017. Production, use, and fate of all plastics ever made. Sci. Adv. 3, e1700782. https://doi.org/10.1126/sciadv.1700782

*Kranzinger, L., Pomberger, R., Schwabl, D., Flachberger, H., Bauer, M., Lehner, M., Hofer, W., 2018. Output-oriented analysis of the wet mechanical processing of polyolefin-rich waste for feedstock recycling. Waste Manag. Res. 36, 445–453

]

---

# Challengue

The development of sustainable industrial supply chains

.left-column[
.footnotesize[
- Only technico-economic criteria alone do not (no longer) enables a informed decision making 
- Need to assess value chains in a different way, including actual impacts on ecosystems
- A more holistic vision is needed to assess the feasibility of these new sectors.
]
 ]

.right-column[ 
.polaroid[ ![](figures/Recycling.png) ]
]

---

## Green Fablab: Towards a distributed recycling via additive manufacturing (DRAM)

.pull-left[

-Local closed loop  <br>
-AM adds value  <br>
-Small quantities   <br>
-Reduction of landfilling  <br>
-Minimize the use of virgin materials  <br>

```{r, include=TRUE, echo=FALSE, fig.cap=" ", out.width = '100%', fig.align='left'}
knitr::include_graphics(paste0(URL, 'Green-Fablab/Green-fablab-2.jpg'))
```

]
.pull-right[

```{r, include=TRUE, echo=FALSE, out.width = '100%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Green-Fablab/Green-fablab-1.jpg'))
```

http://lf2l.fr/projects/green-fablab
]

.footnote[
Cruz Sanchez, F.A., Boudaoud, H., Muller, L., Camargo, M., 2014. Towards a standard experimental protocol for open source additive manufacturing. Virtual Phys. Prototyp. 9, 151–167. https://doi.org/10.1080/17452759.2014.919553
]

---

## Green Fablab: Towards a distributed recycling via additive manufacturing (DRAM)

.left-column[ 
Small scale: 2 km radius (neighbourhood scale) <br>
Low processing capacity, <br>
Technologies: cheap, easy to use <br>
Active role of users <br>

```{r, include=TRUE, echo=FALSE, out.width = '100%', fig.align='center'}
knitr::include_graphics('figures/GF-users.jpg')
```
]

.right-column[ 
.center[![:scale 70%](figures/Green-fablab-circuit.png)]

]

---

class: middle

# Agenda

.pull-left[
1. Who am I?
2. ERPI Laboratory
3. Green Fablab
4. .content-box-yellow[Everest Bio]
5. Proposition and Research axis
]

.pull-right[

```{r, out.width = '100%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Everest-bio/logos/logo-2.jpg'))
```
]

---

## People in Everest Bio

.pull-left[
```{r, out.width = '500px', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Everest-bio/Mauricio.jpg'))
```

```{r, out.width = '500px', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Everest-bio/Hakim.jpg'))
```

```{r, out.width = '500px', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Everest-bio/Fedoua.jpg'))
```

```{r, out.width = '500px', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Everest-bio/Pavlo.jpg'))
```

]

.pull-right[
```{r, out.width = '500px', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Everest-bio/Nadege.jpg'))
```

```{r, out.width = '500px', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Everest-bio/Alexis.jpg'))
```

```{r, out.width = '500px', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Everest-bio/Cristian.jpg'))
```

]

---

## Research Question & Goal

- To develop a generic methodology to assess the **ecosystem services** rendered by industrial activities operating in short circuit in order to improve decision-making by industrial and public actors.

- Validation on the case of a short-circuit plastic recycling chain.

---

## Methodology

- In line with the thesis work previously carried out by the two teams, development of a coupled **Systems Dynamics and LCA approach**.

- Application on a case study, the Green Fablab, by evaluating its environmental impact when setting up a short plastic recycling circuit.

.center[ ![:scale 80%](figures/Goals.png)]

---

## Purpose of Everest Bio

Taking into account the environment and their ecosystem services (ES).

.pull-left[

```{r, out.width = '90%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Ecoservices/Eco-Services-00.jpg'))
```
]

.pull-right[
![](figures/Eco-Services-01.png)

- How the future recycling chain could be including the environment?
]

.footnote[
Bruel, A., Kronenberg, J., Troussier, N., & Guillaume, B. (2019). Linking industrial ecology and ecological economics: A theoretical and empirical foundation for the circular economy. Journal of Industrial Ecology, 23(1), 12-21.
]

---

class: middle

### Ecosystem Services as a **boundary object** for sustainability

.pull-left[

In that sense, Our focus is towards becoming an explicit management tool:

- Creating Systems, Normative and Transformative knowledge using as case study distributed recycling

- better interdisciplinary knowledge integration, greater focus on normative knowledge, and improved recognition of the ecosystem services concept as a potential transformative tool.

]

.pull-right[
```{r, out.width = '90%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'Ecoservices/Abson2014-00.jpg'))
```
]

.footnote[
Abson, D.J., von Wehrden, H., Baumgärtner, S., Fischer, J., Hanspach, J., Härdtle, W., Heinrichs, H., Klein, A.M., Lang, D.J., Martens, P., Walmsley, D., 2014. Ecosystem services as a boundary object for sustainability. Ecol. Econ. https://doi.org/10.1016/j.ecolecon.2014.04.012
]

---

class: top, left
background-image: url(`r paste0(URL, 'Ecoservices/Vanderwilde2021-01.jpg')`)
background-position: 290px 0px
background-size: 75%

### We look to connect <br> Ecosystem services and LCA

Hoping to contributing in <br> sustainability uncertainties.

.footnote[
Vanderwilde, C.P., Newell, J.P., 2021. Ecosystem services and life cycle assessment: A bibliometric review. https://doi.org/10.1016/j.resconrec.2021.105461 <br>
Diwekar, U., Amekudzi-Kennedy, A., Bakshi, B., Baumgartner,et al, 2021. A perspective on the role of uncertainty in sustainability science and engineering. Resour. Conserv. Recycl. 164, 105140. https://doi.org/10.1016/j.resconrec.2020.105140
]

---

### Systems Dynamics of DRAM

Technical system --> Global *system* embedded in a *territory*!

```{r, out.width = '90%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'DRAM/DRAM-Territory-00.jpg'))
```

---

## Towards systemic solutions

Global system

.pull-left[
```{r, out.width = '100%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'DRAM/DRAM-08.png'))
```

]

.pull-right[
- Establish and evaluate scenarios in a systemic and dynamic way
- Evaluate the influences of the main operating variables on the system taking ecosystem services boundaries
- Taking into account the priorities of the multiple stakeholders
- Facilitate your decision making public-private actors
]

.footnote[
Santander, P., Cruz Sanchez, F.A., Boudaoud, H., Camargo, M., 2020. Closed loop supply chain network for local and distributed plastic recycling for 3D printing: a MILP-based optimization approach. Resour. Conserv. Recycl. 154, 104531. https://doi.org/10.1016/j.resconrec.2019.104531 
]

---

# Delivrables & Valorisations

1. An article in an indexed international journal
2. Participation in an international conference 
3. An operational model with a model of a software tool for assessing local impacts
4. Final report on the evaluation of Green Fablab and possible generalisations for short industrial circuits.

- Method and [**prototype of a tool**](https://fabbiocrux.github.io/everest-bio-tool/) for the evaluation of local impacts of circular production systems

---
class: middle, center

# Thanks !

![:scale 70%](figures/DRAM-bd.png)