### 2022-02-14 docs/xar-EB-ES/Questionnaire.Rmd

---
title: "Questionnaire for Everest Bio Tool"
output: pdf_document
header-includes:
   - \usepackage{tabu}
   - \usepackage{booktabs}
   - \usepackage{xcolor}
   - \def\DefaultOptionsofCheckBox{print,bordercolor=blue}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Main Structure of the questionnaire

1. Characterization of the respondent 
   1. Identification of the Geographical area.
   2.
2. Selection of the

## Questionnaire for the Territorial side

\begin{Form}

\begin{tabu} to \linewidth {X[2,l] | X[1,c] | X[1p, c]}

\midrule
\textbf{Characterizing the respondant} & & \\
\midrule

\TextField[width=15em, bordercolor={0 0 0}]{Enter your name:} & & \\

\midrule
\textbf{Provisioning Services (biotic \& Abiotic)} & & \\
\midrule

\textit{Are cultivated plants (Terrestrial or Aquatic) strategic for the territory for nutrition, materials or energy ?} & 
         \CheckBox[name=optionA, width=10pt]{Yes  Not} & \\

 & & \\
 & & \\
 & & \\

\end{tabu}

\end{Form}