### 2022-02-14 docs/xar-EB-general/partials/ERPI.Rmd

---
class: middle

# Agenda

.pull-left[
1. .content-box-yellow[ERPI Laboratory]
1. Green Fablab
1. Everest Bio
1. Proposition and Research axis
]

.pull-right[

.brand-blue[Equipe de Recherche sur le Processus Innovatifs - ERPI]

```{r, out.width = '80%', fig.align='center'}
knitr::include_graphics('https://erpi.univ-lorraine.fr/assets/images/logo-ERPI.svg')
```
]

---

## .small[Equipe de Recherche sur le Processus Innovatifs - ERPI]

A .red[Multidisciplinary] team in **Industrial Engineering** specialized in the study of Innovation processes

.pull-left[

```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics('https://erpi.univ-lorraine.fr/assets/images/logo-ERPI.svg')
```

45 Members average

24 permanent researchers <br>
(10 full Pr., 13 associated Pr, 1 Research engineer) <br>
1 Studies engineer <br>
5 Postdoc  <br>
3 Temporary lecturers  <br>
2 Admin. <br>

]

.pull-right[

```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'LF2L/Methodology-complet-00.jpg'))
```

Research platform: **LF2L** - http://lf2l.fr
]

---

# Research domains at ERPI:

.pull-left[
.pull-left[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/01-Innovation-metrology.jpg'))
```

.orange[
**Innovation capability**

Product <br>
Process  <br>
Business Model <br>

]
]
.pull-right[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/02-Prospective.jpg'))
```
.blue[
**Technological  Emergence/maturity**

Product-Process <br>
Industrial supply chain <br>
]]]

.pull-right[
.pull-left[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/03-Design.jpg'))
```
.green[
**Materialized**

Creative Exploration <br>
Feasibility Product-Process <br>
Physic & digital (virtual) Prototyping <br>
]
]
.pull-right[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/04-Acceptability.jpg'))
```
.purple[
**Front-end validation**
Multi-actors Acceptability <br>
User eXperience & test by use <br>
Living Lab <br>
]]]

.center.Large[Creator of **methods & tools** supporting innovation process]

---

## Scope: The Fuzzy Front End of innovation

.pull-left[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/Scope-00.jpg'))
```
]

.pull-right[
```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/Purpose-00.jpg'))
```
.footnote[
Martinez-Hernandez, E., 2017. Trends in sustainable process design—from molecular to global scales. Curr. Opin. Chem. Eng. 17, 35–41. <br/> https://doi.org/10.1016/j.coche.2017.05.005 
]
]

Think earlier in the sustainable design and their impact

---

## Innovation: a response to current social challenges

```{r, out.width = '95%', fig.align='center'}
knitr::include_graphics(paste0(URL, 'ERPI/Presentation/Fields-00.jpg'))
```

.tiny[Tran Thi Hoang, G., Dupont, L., & Camargo, M. (2019). Application of Decision-Making Methods in Smart City Projects: A Systematic Literature Review. Smart Cities, 2(3), 433-452.]

---